---
title: "Well diffusion"
---

# Packages

```{r}
library(tidyverse)
library(gsheet)
library(cowplot)
library(multcomp)
library(lme4)
library(ggthemes)
library(car)
library(effects)
library(emmeans)
library(afex)  
```

# Load data

```{r}
data1 = gsheet2tbl("https://docs.google.com/spreadsheets/d/1qYZ7EfVSn-gL-pV-bBEVU-LCAdDUAz-SToCaHQpZjlE/edit?usp=sharing") %>% 
  mutate(long = as.numeric(long),
         trans = as.numeric(trans)) %>% 
  mutate(dai = case_when(aval == 1 ~ 5,
                         aval == 2 & species=="microtom" & trial==1 ~ 10,
                         aval == 2 & species=="microtom" & trial==2 ~ 7,
                         aval == 2 & species=="potato" ~ 7,
                         aval == 3 ~9,
                         aval == 4 ~11)) %>% 
  mutate(leasion_size = (trans+long)/2) %>% 
  filter(leasion_size>0) %>% 
  mutate(trat = factor(trat, levels = c("CHECK","MEDIUM","CE_MEDIUM","CE_RESIDUE", "CE_LPP")))

head(data1)
```

# Visualization

```{r}
data1 %>% 
  ggplot(aes(dai,sqrt(leasion_size), color = trat))+
  geom_point()+
  geom_smooth(method = "lm", se = F, formula = y~poly(x, 1))+
  facet_grid(species ~trial)

```

```{r}
data1 %>% 
  group_by(trial,trat,species) %>% 
  filter(dai == max(dai)) %>% 
  ggplot(aes(trat,leasion_size))+
  geom_jitter(width = 0.1)+
  geom_boxplot(fill = NA)+
  facet_grid(species ~trial)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))
```

# Microtom

Selecting microtom tom data

```{r}
data_microtom = data1 %>% 
  filter(species == "microtom") %>% 
  mutate(leaflet = as.factor(leaflet),
         trial = as.factor(trial)) %>% 
  mutate(trat = factor(trat, levels = c("CHECK","MEDIUM","CE_MEDIUM","CE_RESIDUE", "CE_LPP")))


```

# modeling

## Mix model

```{r}
mix_microtom = lmer(sqrt(leasion_size) ~ trat*dai + (1|trial/leaflet),
                    data = data_microtom, REML = F)
summary(mix_microtom)
AIC(mix_microtom)
```

```{r}
Anova(mix_microtom)
```

```{r}
# em1 = emtrends(mix_microtom,~trat, trend = "dai")
```

## Visualize

```{r}
eff = predictorEffects(mix_microtom, "trat", xlevels = 50)
# eff$trat
as.data.frame(eff$trat) %>%
  mutate(trat = factor(trat, levels = c("CHECK","MEDIUM","CE_MEDIUM","CE_RESIDUE", "CE_LPP"))) %>% 
  # mutate(fit = exp(fit),
         # lower = exp(lower),
         # upper = exp(upper) ) %>% 
  ggplot(aes(dai,fit,color =trat))+
  geom_point(data =data_microtom, aes(dai,sqrt(leasion_size), group=trat),color = "gray")+
  geom_ribbon(aes(ymin =(lower),ymax = (upper), group = trat),color =NA, alpha = 0.01)+
  geom_line(size = 1)+
  scale_color_colorblind()+
  theme_minimal_hgrid()+
  labs(x = "dai",
       y = expression(sqrt("diameter")))
  # facet_wrap(~trat)

```

## Final severity

```{r}
microton_final = data1 %>% 
  filter(species =="microtom") %>% 
  group_by(trial,trat,species) %>% 
  filter(dai == max(dai)) 

```

```{r}
mix_microtom2 = lmer((leasion_size) ~ trat + (1|trial/leaflet),
                    data = microton_final, REML = F)
summary(mix_microtom2)
AIC(mix_microtom2)
```

# Well difusion

## Load data

```{r}
well_load = gsheet2tbl("https://docs.google.com/spreadsheets/d/1rFebMOmrogBPB3JSVTKfDtQIFU_B6pnwseSJPWcLVt4/edit?usp=sharing")
```

# Tidy

### Lote 1

```{r}
well_data1 = well_load %>% 
  mutate(trat =  factor(trat, levels = c("Check" ,"CE_Medium", "B157", "CE_Lpp","Frac1","Frac2","Frac3","Frac4","Frac5","Frac6","Frac7","Frac8","Frac9","Frac10","Frac11")))%>% 
  filter(lote ==1)

```

## viz

```{r}
well_data1  %>% 
  ggplot(aes(trat, raio, color = as.factor(exec), fill =as.factor(exec)))+
  geom_point(color ="gray",
             size =1,
             position = position_jitterdodge(jitter.width = .3,
                                             dodge.width = .5))+
  stat_summary(fun.data = "mean_cl_boot",
               geom = "errorbar",
               width = 0,
               size =.5,
               position = position_dodge(width=.5))+
  stat_summary(fun = "mean",
               size = 1.4,
               geom = "point",
               position = position_dodge(width=.5))+
  facet_wrap(~species)+
  scale_color_calc()+
  theme_minimal_hgrid(font_size = 10)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        panel.border = element_rect(color = "gray"),
        strip.text = element_text(face = "italic"))+
  labs(x = "",
       y = "Inhibition zone (radius cm)",
       color = "Trial",
       fill = "Trial")


```

## Modeling

### trial 1

```{r}
t1_well_data1 = well_data1 %>% 
  filter(exec ==1)
```

```{r}
leveneTest(raio ~ trat*species, t1_well_data1)
```

Inhibition zone values equal to zero were converted to 0.001 because the Gamma distribution assumes only positive values

```{r}
mod1 = glm(raio+0.001 ~ trat*species,family = Gamma, t1_well_data1)
summary(mod1)
# hist(residuals(mod1))
# shapiro.test(residuals(mod1))
```

```{r}
Anova(mod1)
```

```{r}
cld(emmeans(mod1, "trat", by = "species", type= "response"), Letters = letters)
plot
```

```{r}
# cld(emmeans(mod1, "species", by = "trat", type= "response"), Letters = letters)
```

### Trial 2

```{r}
t2_well_data1 = well_data1 %>% 
  filter(exec ==2)
```

Inhibition zone values equal to zero were converted to 0.001 because the Gamma distribution assumes only positive values

```{r}
mod2 = glm(raio+0.001 ~ trat*species,family = Gamma, t2_well_data1)
summary(mod2)

```

```{r}
Anova(mod2)
```

```{r}
cld(emmeans(mod2, "trat", by = "species", type= "response"), Letters = letters)
```

```{r}

# cld(emmeans(mod2, "species", by = "trat"), Letters = letters)
```

### Lote 2

```{r}
order = c("Check" ,"CE_Medium", "CP", "CE_Lpp","Frac1","Frac2","Frac3","Frac4","Frac5","Frac6","Frac7","Frac8","Frac9","Frac10","Frac11")
well_data2 = well_load %>% 
  mutate(trat =  factor(trat, levels = c("Check" ,"CE_Medium", "B157", "CE_Lpp","Frac1","Frac2","Frac3","Frac4","Frac5","Frac6","Frac7","Frac8","Frac9","Frac10","Frac11")))%>% 
  filter(lote ==2, trat !="Frac8") %>%
  mutate(trat = as.character(trat),
         trat = case_when(trat == "Frac9" ~ "Frac8", trat == "Frac10" ~ "Frac9",
                          !trat %in% c("Frac9", "Frac10") ~ trat)) %>%
  mutate(trat =  factor(trat, levels =order ))

```

## viz

```{r}
well_data2  %>% 
     ggplot(aes(trat, raio, color = as.factor(exec), fill =as.factor(exec)))+
  geom_point(color ="gray",
             size =1,
             position = position_jitterdodge(jitter.width = .3,
                                             dodge.width = .5))+
  stat_summary(fun.data = "mean_cl_boot",
               geom = "errorbar",
               width = 0,
               size =.5,
               position = position_dodge(width=.5))+
  stat_summary(fun = "mean",
               size = 1.4,
               geom = "point",
               position = position_dodge(width=.5))+
  facet_wrap(~species)+
  scale_color_calc()+
  theme_minimal_hgrid(font_size = 10)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        panel.border = element_rect(color = "gray"),
        strip.text = element_text(face = "italic"))+
  labs(x = "",
       y = "Inhibition zone (radius cm)",
       color = "Trial",
       fill = "Trial")

```

## Modeling

### trial 1

```{r}
t1_well_data2 = well_data2 %>% 
  filter(exec ==1)
```

```{r}
mod1_2 = glm(raio+0.001 ~ trat*species,family = Gamma, t1_well_data2)
summary(mod1_2)
```

```{r}
boot::glm.diag.plots(mod1_2,boot::glm.diag(mod1_2))

```

```{r}
Anova(mod1_2)
```

```{r}
cld(emmeans(mod1_2, "trat", by = "species", type= "response"), Letters = letters)
plot
```

```{r}
# cld(emmeans(mod1, "species", by = "trat", type= "response"), Letters = letters)
```

### Trial 2

```{r}
t2_well_data2 = well_data2 %>% 
  filter(exec ==2)
```

```{r}
mod2_2= glm(raio+0.001 ~ trat*species, family = Gamma, t2_well_data2)
summary(mod2_2)

```

```{r}
boot::glm.diag.plots(mod1_2,boot::glm.diag(mod2_2))

```

```{r}
Anova(mod2_2)
```

```{r}
cld(emmeans(mod2_2, "trat", by = "species", type= "response"), Letters = letters)
```

```{r}

# cld(emmeans(mod2, "species", by = "trat"), Letters = letters)
```

```{r}
all_data_modeled = bind_rows(
as.data.frame(emmeans(mod1, "trat", by = "species", type= "response")) %>% 
  mutate(lote = "Second batch",
         trial = "First trial"),
as.data.frame(emmeans(mod2, "trat", by = "species", type= "response"))%>% 
  mutate(lote = "Second batch",
         trial = "Second trial"),

as.data.frame(emmeans(mod1_2, "trat", by = "species", type= "response")) %>% 
  mutate(lote = "Third batch",
         trial = "First trial"),
as.data.frame(emmeans(mod2_2, "trat", by = "species", type= "response"))%>% 
  mutate(lote = "Third batch",
         trial = "Second trial")


)


```

```{r}
all_data_modeled %>%
  filter(lote == "Second batch") %>% 
    mutate(trat = as.character(trat),
         trat = case_when(trat == "B157" ~ "CP",
                          trat == "CE_Medium" ~ "CE medium",
                          trat == "Check"~ "Control",
                          trat == "CE_Lpp" ~ "CE",
                          !trat %in% c("B157","CE_Medium", "Check", "CE_Lpp") ~ trat),
         species = case_when(species == "Isolate A3 - A. linariae" ~ "A3 - A. linariae",
                            species == "Isolate 248 - A. grandis" ~ "As248P - A. linariae")) %>%
    mutate(trat =  factor(trat, levels = c("Control" ,"CE medium", "CP", "CE","Frac1","Frac2","Frac3","Frac4","Frac5","Frac6","Frac7","Frac8","Frac9","Frac10","Frac11")))%>% 
   ggplot(aes(trat,response, color = trial))+
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                position = position_dodge(width = 0.4),
                width = 0, size =0.5)+
  geom_point(position = position_dodge(width = 0.4), size = 1.5)+
  
  lemon::facet_rep_wrap(~species)+
  scale_color_manual(values= c("#41ae76","#005824"))+
  theme_half_open(font_size = 12, line_size = 0.2)+
  theme(
        strip.text = element_text(face = "italic"),
        legend.position = "top",
        strip.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)
        )+
  labs(x = "",
       y = "Inhibition zone (cm)",
       color ="")+
  background_grid(size.major = 0.2,color.major = "grey90")
  # coord_flip()
ggsave(".figs/inhib.png", dpi = 600, height = 3, width = 7, bg = "white")


```
